---
title: "lab3_script"
author: "Wanjia Guo"
date: "2/6/2021"
output: 
  html_document
---

```{r setup, include=FALSE}
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(gghighlight)
library(colorblindr)

df <- read_csv(here::here("data", "lab3", "honeyproduction.csv"))
```

### Visualize the total production of honey across years by state Use color to highlight the west coast (Washington, Oregon, and California) with a different color used for each west coast state.


```{r}

p1 <- ggplot(df, aes(year, totalprod)) + 
  geom_line(aes(group=state),
            color="lightgrey") + 
  geom_line(data = df %>% filter(state %in% c("CA", "OR", "WA")), 
            aes(color=state), 
            size=1.5) + 
  theme_minimal()

p1

```

### Reproduce the plot according three different kinds of color blindness, as well as a desaturated version.

```{r}

cvd_grid(p1)

```

### Reproduce the plot using a color blind safe palette.

```{r}

p2 = p1 + scale_color_viridis_d()

p2

cvd_grid(p2)

```

### Download the file here denoting the region and division of each state.
#### Join the file with your honey file.

```{r}

df2 <- read_csv(here::here("data", "lab3", "regions_divisions.txt"))

df3 <- df %>% left_join(df2, by=c("state"="State Code"))

```

#### Produce a bar plot displaying the average honey for each state (collapsing across years).

```{r}
df_avg = df3 %>% 
  group_by(Region, state) %>% 
  summarise(year_m = mean(totalprod))

ggplot(df_avg, aes(x=state, y=year_m)) + 
  geom_bar(stat="identity") + 
  coord_flip() + 
  theme_minimal()
```

#### Use color to highlight the region of the country the state is from.
Note patterns you notice: The production of honey is an exponential distribution, and most of tail are Northeast and South.

```{r}
ggplot(df_avg, aes(x=reorder(state, year_m), 
                   y=year_m, 
                   fill=Region)) + 
  geom_bar(stat="identity") + 
  coord_flip() + 
  scale_fill_OkabeIto() +
  labs(x="State", 
       y="Total production (averaged over years)") + 
  theme_minimal()
```

### Create a heatmap displaying the average honey production across years by region (averaging across states within region).

```{r}

df_avg2 = df3 %>% 
  group_by(Region, year) %>% 
  summarise(year_m = mean(totalprod))

ggplot(df_avg2, aes(x=year, 
                    y=Region, 
                    fill=year_m)) + 
  geom_tile() + 
  scale_fill_viridis_c() + 
  labs(fill = "Production") + 
  coord_fixed() + 
  theme_minimal() 

```


### Create at least one more plot of your choosing using color to distinguish, represent data values, or highlight.

I tried for a while but had difficulties in installing the "albersusa" package, which seems to be caused by a failure to install the "units" package. When I tried to install the "unit" manually, the error message says I need the "libudunits2.so" to install the "units", so I tried to install udunits through command line. Then the command line says it was installed already, so I am a bit run out of ideas for the installation. 

```{r}
df_avg3 = df3 %>% 
  group_by(state) %>% 
  summarise(m=mean(priceperlb))

ggplot(df3, aes(year, 
                priceperlb, 
                color=state)) + 
  geom_line(aes(group=state)) + 
  facet_wrap(~state)+
  gghighlight(max(priceperlb) > 3.5, 
              use_direct_label=FALSE) + 
  scale_color_OkabeIto(darken = 0.2) +
  theme_minimal() +
  labs(title = "States with honey unit price over $3.5 over time", 
       y = "price per lb") + 
  theme(legend.position = "none")

```

